Type-Safe Modular Parsing

### 2017-06-05 SLE_17/sections/CaseStudy.tex
Shorten conclusion. Resize table.
Fit into 12 pages
Revision

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{resources/depGraph.pdf}
    \caption{Dependency graph of all calculi and components. Grey boxes are calculi; white boxes are components.}
    \label{fig:dependency}
\end{figure*}

@@

From the graph, we know that common components such as \lstinline{VarApp} are reused in
lots of calculi. Such reuse could shorten the code considerably.
Next subsection examines amount of code as well as performance.

@@

\begin{table*}
    \centering
    \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|}
      \hline
        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{7}{ c| }{\bfseries Time (ms)} \\ \cline{2-11}
        \multicolumn{1}{|c|}{} & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyalongest{} & \bfseries (+/-)\% & \ourclass{} & \bfseries (+/-)\% \\
      \hline
      \input{resources/CaseStudyTable.tex}
      \hline
      \multicolumn{7}{c}{}
    \end{tabular}
    \caption{Comparison of SLOC and execution time.}
    \label{tab:comparison}
\end{table*}

@@

To discover the reasons of slower execution time, we made experiments
on two possible factors, which are Object Algebra and the longest match alternative combinator.
We use Object Algebra for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
while \ilyaimpl{} uses case class and the ordinary alternative combinator.
Therefore, we implemented two more versions. One is a modified version of our implementation,
named \ourclass{}, with object algebra replaced by case class for the ASTs.
The other is a modified version of \ilyaimpl{}, named \ilyalongest{},
using the longest match alternative combinator instead of the ordinary one.

@@

The right part of table \ref{tab:comparison} suggests that the difference of running time between
using object algebra and class is little, roughly 1\%.
The usage of longest match combinator slows the performance by 7\%. The main reason of slower
execution time may be the overall structure of the modular parsing approach, because we indeed have
more intermediate function calls and method overriding. However, it is worth mentioning that
because of the memoization technique of Packrat parsers, we are only constant times
slower, the algorithmic complexity is still the same.

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{resources/depGraph.pdf}
    \caption{Dependency graph of all calculi and components. Grey boxes are calculi; white boxes are components.}
    \label{fig:dependency}
\end{figure*}

@@

%% From the graph, we know that common components such as \lstinline{VarApp} are reused in
%% lots of calculi. Such reuse could shorten the code considerably.
%% Next subsection examines amount of code as well as performance.

@@

\begin{table*}
    \centering
    \resizebox{0.7\textheight}{!}{
    \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|}
      \hline
        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{7}{ c| }{\bfseries Time (ms)} \\ \cline{2-11}
        \multicolumn{1}{|c|}{} & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyalongest{} & \bfseries (+/-)\% & \ourclass{} & \bfseries (+/-)\% \\
      \hline
      \input{resources/CaseStudyTable.tex}
      \hline
      \multicolumn{7}{c}{}
    \end{tabular}}
    \caption{Comparison of SLOC and execution time.}
    \label{tab:comparison}
    \vspace{-0.3cm}
\end{table*}

@@

To discover the reasons of slower execution time, we made experiments
on two possible factors, which are Object Algebras and the longest match alternative combinator.
We use Object Algebras for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
while \ilyaimpl{} uses case class and the ordinary alternative combinator.
Therefore, we implemented two more versions. One is a modified version of our implementation,
named \ourclass{}, with Object Algebras replaced by case class for the ASTs.
The other is a modified version of \ilyaimpl{}, named \ilyalongest{},
using the longest match alternative combinator instead of the ordinary one.

@@

The right part of table \ref{tab:comparison} suggests that the difference of running time between
using Object Algebras and class is little, roughly 1\%.
The usage of longest match combinator slows the performance by 7\%. The main reason of slower
execution time may be the overall structure of the modular parsing approach, because we indeed have
more intermediate function calls and method overriding. However, it is worth mentioning that
because of the memoization technique of Packrat parsers, we are only constant times
slower, the algorithmic complexity is still the same.

### 2017-06-05 SLE_17/sections/Discussion.tex
Revision

Our prototype uses Packrat parsing as the underlying parsing technique, OO inheritance for composing and extending parsers, and Object Algebras for parsing extensible ASTs. Yet such a framework is itself flexible and modular, because those techniques can have alternatives.
For example, as we mentioned before, any parsing library
that resolves the algorithmic challenges in modular parsing can work well. Regarding OO inheritance for the extensibility, an alternative approach, called \textit{open recursion}~\cite{CookThesis} can be used in other languages, by introducing explicit ``self-reference'' parameters for the recursion. Furthermore, besides Object Algebras, v\textit{Data types à la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures. For the goal of modular parsing a custom combination of those alternatives can be adopted.

>>>>

Our prototype uses Packrat parsing as the underlying parsing technique, OO inheritance for composing and extending parsers, and Object Algebras for parsing extensible ASTs. Yet such a framework is itself flexible and modular, because those techniques can have alternatives.
For example, as we mentioned before, any parsing library
that resolves the algorithmic challenges in modular parsing can work well. Regarding OO inheritance for the extensibility, an alternative approach, called \textit{open recursion}~\cite{CookThesis} can be used in other languages, by introducing explicit ``self-reference'' parameters for the recursion. Furthermore, besides Object Algebras, \textit{Data types \`a la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures. For the goal of modular parsing a custom combination of those alternatives can be adopted.
